Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-06 cluster_paper/ms.tex
update ms

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be tidal escapees from the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  sensitive clustering algorithms.
  In this study, we focus on a halo with a tip-to-tip length of
  $\approx$500\,pc reported around the $\approx$150\,Myr open cluster
  NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two thirds of kinematically selected halo members out to 250\,pc from the
  cluster center have rotation periods consistent with a
  gyrochronological age of 150\,Myr.
  A comparison sample of field stars shows no such trend.
  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations.
  This work expands the set of confirmed NGC 2516 members by a factor
  of two, and is consistent with a paradigm in which the halos of many
  open clusters are more populous than their cores.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

One point of difficulty however is that different clustering methods
yield differing levels of both sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.},
\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
Differences between different clustering methods raise the issue of
how contaminated the resulting samples are, and in some cases whether
the reported structures are truly coeval stellar populations at all.

@@

As part of this broader project, we focus here on a modest question:
in just a single southern open cluster, is the cluster halo truly
coeval with the core?  We chose NGC\,2516 for this analysis since it
is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
to facilitate photometric rotation measurements using TESS, and some
degree of spectroscopic analysis.  Its halo is decidedly
non-spherical: \citet{kounkel_untangling_2019} reported its dimensions
as $\approx20\times10\times350$\,pc.  We want to know: is the halo
real? To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of field stars?  What are the
implications, observationally and theoretically, if we can identify
the dispersed halos of open clusters?

%TODO: double check this owrding.
A brief note on terminology.  Low-density stellar associations
connected to a dense population (the ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'',
``outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
association.  We opt for ``halo'' because it is concise and model
agnostic.  It suffers however from the suggestion of spherical
symmetry.  The halo of NGC\,2516 might be best described as ``tails'',
but the primary goal of this study is not to verify whether these
tails are tidal.

Section~\ref{sec:gaia} summarizes the astrometric analyses of the Gaia
data that led to our interest in NGC\,2516.  Section~\ref{sec:agedate}
age-dates the cluster's halo and core using Gaia photometry
(Section~\ref{subsec:hr}), TESS gyrochronology
(Section~\ref{subsec:tess}), and lithium depletion
(Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
discuss the implications of our analysis for NGC\,2516 specifically
and stellar spin-down and open cluster evolution generally.
Section~\ref{sec:conclusion} gives our conclusions.

@@

%\subsection{Gaia Astrometry}
%\label{subsec:astrometry}

@@

We selected candidate \cn\ members based on those reported in the
literature.  While some pre-Gaia analyses were available, the purity
and accuracy of the Gaia-derived results are the current state of the
art \citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}.  We therefore
adopted what we viewed as the most interesting clustering samples to
compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of the different samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}.}.  While we considered performing our
own clustering analysis using the Gaia data, such an effort would in
effect only replicate the work of these investigators.  We opt to
instead use their studies as starting points.

The Gaia clustering studies each used different selection functions,
which yielded different results.  \citetalias{cantatgaudin_gaia_2018}
considered stars brighter than $G=18$\,mag within a few degrees of the
cluster center, and reported 1106 candidate \cn\ members.
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
considered stars up to $\approx$1\,mag fainter, and reported 3003 and
1860 members respectively.  The unsupervised clustering techniques
that each of these studies applied to the second Gaia data release are
discussed Appendix~\ref{app:clustering}, as is the overlap between
their resulting membership samples.

Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study in the space of observed positions, proper motions, and radial
velocity when available.  The \citetalias{cantatgaudin_gaia_2018}
members are all within a few degrees of the cluster center, while the
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members (the ``halo'') span tens of degrees.  Inverting the parallax
shows that members have been reported from line-of-sight distances
ranging from $300$ to $600\,$pc.  Stars in the upper right quadrant of
the $\mu_{\alpha'}$ {\it vs.} $\mu_\delta$ projection were reported
almost entirely by \citet{meingast_2021}.

In Figure~\ref{fig:gaia6d}, and for the remainder of the study, we
describe the \citetalias{cantatgaudin_gaia_2018} members as the
``core'' of the cluster, and any non-overlapping
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members as the ``halo''.  This defintion implies that there are
\ncore\ core members, and \nhalo\ halo members.  The distinction is
perhaps tautological, because \citetalias{cantatgaudin_gaia_2018} did
not extend their search for members out to tens of degrees from the
cluster center.  Nonetheless, the \citetalias{cantatgaudin_gaia_2018}
membership catalog is consistent with that of many previous
investigators \citep[{\it
e.g.},][]{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, and is
consistent with the general {\it visual} impression that one has when
looking at \cn\ in the night sky: it appears to span
$\approx$2$^\circ$, not $\approx20^\circ$.

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following intervals of right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
allowed the number of stars in the comparison sample to exceed that in
the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.  We also required the
comparison sample to not overlap with the cluster sample, which led to
the omission of 1.1\% of the stars drawn from the volume noted above.

The discrepancy between our canonical understanding of open clusters
and the possibility of a diffuse population existing around \cn\
prompted our interest in understanding the reality and origin of the
diffuse population.  Figure~\ref{fig:XYZ} helps in visualizing the
proposed structure and kinematics of the halo.  We computed the
physical positions by transforming from $(\alpha, \delta, \pi)$ to
galactocentric ($X,Y,Z$) assuming the \texttt{astropy v4.0} coordinate
standard \citep{astropy_2018}.  The direction of galactic rotation is
$+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
median parameters for core members for which
\citetalias{cantatgaudin_gaia_2018} reported membership probabilities
exceeding 70\%.  We then integrated the orbit using \texttt{gala} and
the \texttt{MWPotential2014} potential \citep{bovy_galpy_2015,gala}.
The elongation of the cluster in both the $(X,Y)$ and $(Z,Y)$ planes
is correlated with the direction of the LSR-corrected median cluster
velocity.  A similar correlation was noted in Coma Ber by
\citet{tang_comaber_2019}.  Possible explanations are discussed in
Section~\ref{subsec:origin}.

@@

The first check on the plausibility of the candidate cluster members
being coeval is to analyze the Gaia Hertzsprung-Russell (HR) diagrams.
Comparable analyses have already been performed by
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $Bp-Rp$ color.  The core
members of the cluster show a clean sequence consistent with stars
with a fixed age and metallicity, and varying mass.  The halo members
show a consistent sequence, though with greater scatter.  One possible
explanation for the additional scatter could be that the halo is more
contaminated by field stars.  For instance, $\approx5$ halo
stars are present in portions of the red giant branch
(RGB) that imply stellar ages inconsistent with that of the cluster.
These stars must be field interlopers.

@@

To determine the reddening correction across the NGC\,2516 cluster, we
queried the 2MASS
DUST
service\footnote{
	\url{http://irsa.ipac.caltech.edu/applications/DUST};
	query performed using the \texttt{astrobase.services.dust} module
	\citep{bhatti_astrobase_2018}.
}
and retrieved the extinction parameters at the positions of each
NGC\,2516 member.  The mean and standard
deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}: $E(B-V)_{\rm
S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
the \citet{schlegel_maps_1998} maps overestimate the reddening values
when the color excess exceeds about 0.10\,mag. Applying the correction
proposed by \citet{bonifacio_search_2000} leads to our adopted
$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
which is used in the plots.  To redden the isochrones, we assumed
$R_V=3.1$, and applied the \citet{odonnell_1994} SED-dependent
extinction law star-by-star through the PARSEC interface.

@@

Overall, the data and models agree for masses above approximately
0.45\,$M_\odot$.  Below this mass, the data and models diverge at
colors redder than $Bp-Rp\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.\footnote{The 100 Myr isochrone does overlap well with the
lower main sequence, but it fails to fit the upper main sequence and
red giants.}  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Proposed explanations for the discrepancy
between the models and observations have include starspots and
incomplete molecular line lists\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
Ultimately, we adopted the PARSEC isochrones because
\citet{chen_improving_2014} implemented an empirical
temperature-opacity calibration, which leads the PARSEC isochrones to
diverge at slightly lower mass than the MIST models.  Regardless, for
purposes of age-dating the cluster we focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal
predictive power.

\citet{cummings_2018} presented techniques for mitigating some of the
complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
rotation, high binarity fractions, and the presence of blue
stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
cluster memberships, they found MSTO ages for NGC\,2516 ranging from
165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
PARSEC, MIST, or SYCLIST).  Their statistical uncertainties ($\approx
25$\,Myr) were comparable to the systematic uncertainties.

Our primary goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} shows that
isochronally, they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim
can be made.  The MSTO stars are consistent with the 178 and 316 Myr
models, but are bluer than the 100 Myr model.  The RGB stars are
consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
model.  Based on the assumption that the width of the MSTO can be
attributed to binary stars, we are most interested in the blue edge
(blue stragglers are, however, a concern).  The blue edge appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for a more precise model-dependent
comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
adopt an PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
our quoted uncertainty is a quadrature sum of their quoted statistical
and systematic components.  The relative age of ``slightly older than
the Pleiades'' is likely the more robust statement, consistent with
reports since \citet{mermilliod_comparative_1981}.

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
    columns}) undergo successive degrees of automated cleaning,
    described in Section~\ref{subsubsec:cluster}.  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
    which some NGC\,2516 members appear slightly elevated with respect
    to the Pleiades, suggesting an older age.  
    Stars redder than $(Bp-Rp)_0\gtrsim 2.2$ (later than M2V) 
    are fainter than the magnitude limit of our survey.
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference is due to 34 stars unique to
\citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and a few stars falling on the TESS chip gaps.  The
$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
the distance of \cn.

We removed systematic trends shared between all light curves on each
CCD in each sector individually, and stitched together the resulting
light curves before searching for rotation-induced periodicity.
Details regarding our detrending approach are presented in
Appendix~\ref{app:detrending}.  After applying an initial detrending
step aimed at removing systematic trends, we proceeding with a few
small cleaning steps to improve the purity of the rotation period
measurements: we masked 0.7 days at the beginning and end of each
spacecraft orbit, and ran a sliding standard-deviation rejection
window over the light curve, which removed any outlying points within
$\pm3\times$MAD of the median in each window.

>>>>

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  It is
  important to verify that these ``halos'', ``tails'', and ``strings''
  are of similar ages and compositions as stars in the denser part of
  the cluster.  We present an analysis of NGC\,2516
  ($\approx$150\,Myr), which has a classical tidal radius of 10\,pc
  and an apparent halo of stars spanning 500\,pc.  Combining
  photometry from Gaia, rotation periods from TESS, and lithium
  measurements from Gaia-ESO and GALAH, we find that the halo of
  NGC\,2516 is real and coeval with the core.  Two thirds of
  kinematically selected halo members out to 250\,pc from the cluster
  center have rotation periods consistent with a gyrochronological age
  of 150\,Myr.  A comparison sample of field stars shows no such
  trend.  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations of
  the Pleiades and Psc\,Eri.  Broadly speaking, this work advances the
  paradigm in which the halos of many open clusters are more populous
  than their cores.  We highlight the implications for spectroscopic
  survey targeting, open cluster dispersal, and planet searches around
  young stars.
\end{abstract}

@@

One point of difficulty however is that different clustering methods
yield differing levels of sensitivity and precision
\citep{hunt_clustering_2020}.  Using say a Gaussian Mixture Model
tautologically yields open clusters that are elliptical \citep[{\it
e.g.},][]{wallace_m4_2018}.  Some unsupervised methods yield dispersed
and asymmetric structures with number densities down to 100 times less
than the field around the same regions (see {\it e.g.},
\citealt{kounkel_untangling_2019} and \citealt{meingast_2021}).
Differences between different clustering methods raise the issue of
how contaminated the resulting samples are, and in some cases whether
the reported structures are truly coeval stellar populations at all.

@@

As part of this broader project, we focus here on a modest question:
in just a single southern open cluster, is the cluster halo truly
coeval with the core?  We chose NGC\,2516 for this analysis since it
is young (100--200\,Myr) and sufficiently nearby ($d\approx400$\,pc)
to facilitate photometric rotation measurements using TESS.  Its halo
was recently reported by \citet{kounkel_untangling_2019} to span
$\approx20\times10\times350$\,pc.  We want to know: is this halo real?
To what extent can we use Gaia alone to reliably identify age-dated
needles in the haystack of field stars?  What are the implications,
observationally and theoretically, if we can identify the dispersed
halos of open clusters?

A brief note on terminology.  Low-density stellar associations
connected to a dense population (a ``core'') have been described as
being in ``halos'', ``strings'', ``coronae'', ``snakes'',
``outskirts'', and ``tidal tails'' \citep[{\it
e.g.},][]{davenport_death_2010,kounkel_untangling_2019,roser_hyades_2019,tian_discovery_2020,meingast_2021}.
The latter term implies a particular model for the formation of the
dispersed group.  ``Halo'' is concise and model agnostic, but suffers
from the suggestion of spherical symmetry.  The halo of NGC\,2516 is
most accurately described as consisting of a leading and trailing
tail. However, the goal of this study is to clarify whether the stars
in these tails are the same age as stars in the core, and not to make
a ruling on the origin of the overall structure.  For this reason, we
adopt the model-agnostic ``halo''.

Section~\ref{sec:gaia} summarizes the astrometric analyses of the Gaia
data that led to our interest in \cn.  Section~\ref{sec:agedate}
measures the age of the cluster's halo and core using Gaia photometry
(Section~\ref{subsec:hr}), TESS gyrochronology
(Section~\ref{subsec:tess}), and lithium depletion
(Section~\ref{subsec:lithium}).  In Section~\ref{sec:discussion} we
discuss the implications of our analysis for NGC\,2516 specifically
and stellar spin-down and open cluster evolution generally.
Section~\ref{sec:conclusion} gives our conclusions.

@@

@@

We selected candidate \cn\ members based on those reported in the
literature.  While some pre-Gaia analyses were available
\citep{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, the purity and
accuracy of the Gaia-derived results are the current state of the art.
We therefore adopted what we viewed as the most important clustering
samples to compare: those of \citet{cantatgaudin_gaia_2018},
\citet{kounkel_untangling_2019} and \citet{meingast_2021}.  We refer
to these studies in the following as
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}
respectively.  A useful visualization of these samples is
available online\footnote{
  \url{https://homepage.univie.ac.at/stefan.meingast/coronae.html},
  made by \citet{meingast_2021}, last accessed \texttt{2021/04/05}.}.
While we considered performing our own clustering analysis on the
Gaia data, such an effort would in effect only replicate the work of
these investigators.  We opt to instead use their studies as a starting
point.

The Gaia clustering studies each used different selection functions,
which yielded different results.  \citetalias{cantatgaudin_gaia_2018}
considered stars brighter than $G=18$\,mag within a few degrees of the
center of \cn, and reported 1106 candidate cluster members.
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
considered stars up to $\approx$1\,mag fainter, and reported 3003 and
1860 members respectively.  The unsupervised clustering techniques
that each of these studies applied to the second Gaia data release are
discussed in Appendix~\ref{app:clustering}, as is the overlap between
their resulting membership catalogs.

Figure~\ref{fig:gaia6d} shows the cluster members reported by each
study in the space of observed positions, proper motions, and radial
velocity when available.
In Figure~\ref{fig:gaia6d}, and for the remainder of the study, we
describe the \citetalias{cantatgaudin_gaia_2018} members as the
``core'' of the cluster, and any non-overlapping
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
members as the ``halo''.  This defintion implies that there are
\ncore\ core members, and \nhalo\ halo members.  The distinction is
tautological, in the sense that \citetalias{cantatgaudin_gaia_2018} did
not extend their search for members out to tens of degrees from the
cluster center.  Nonetheless, the \citetalias{cantatgaudin_gaia_2018}
membership catalog is consistent with that of many earlier
investigators \citep[{\it
e.g.},][]{jeffries_ngc2516_2001,Kharchenko_et_al_2013}, and is
consistent with the general visual overdensity one sees when viewing
\cn\ in the sky: it appears to span $\approx$2$^\circ$, not
$\approx20^\circ$.

Outside of the core and halo, we also define a set of nearby field
stars in the ``neighborhood'' of \cn.  Based on the observed
distribution of halo members, we drew these stars randomly from the
following intervals of right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries using
the \texttt{astroquery.gaia} module \citep{astroquery_2018}.  We
allowed the number of stars in the comparison sample to exceed that in
the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.  We also required the
comparison sample to not overlap with the cluster sample.%, which led to
%the omission of 1.1\% of the stars drawn from the volume noted above.

The style of visualization given in Figure~\ref{fig:gaia6d} does not
make it clear, in our eyes, why the clustering algorithms have decided
to associate certain stars with the halo.  Canonically, open clusters
are spherical, and span $\approx$10\,pc \citep[{\it
e.g.},][]{Kharchenko_et_al_2013}.  Inverting the parallaxes in
Figure~\ref{fig:gaia6d} shows that members have been reported from
line-of-sight distances ranging from $\approx$$300$ to
$\approx$$600\,$pc. 
Is this structure real? What explains its existence?

% Are these halo stars truly associated with the cluster's core?
% What is the 
% The opportunity of verifying and understanding the diffuse
% population of \cn\ prompted our interest in understanding
% the reality and origin of the halo.

%Stars in the upper right quadrant of the $\mu_{\alpha'}$ {\it vs.}
%$\mu_\delta$ projection were reported almost entirely by
%\citet{meingast_2021}.

An initial step in visualizing the reported structure and kinematics
of the candidate cluster members is to consider them in Cartesian
coordinates (Figure~\ref{fig:XYZ}).  We computed the physical
positions by transforming from $(\alpha, \delta, \pi)$ to
galactocentric ($X,Y,Z$) assuming the \texttt{astropy v4.0} coordinate
standard \citep{astropy_2018}.  The direction of galactic rotation is
$+\hat{Y}$. The cluster orbit (gray line) was evaluted by taking the
median parameters for core members for which
\citetalias{cantatgaudin_gaia_2018} reported membership probabilities
exceeding 70\%.  We then integrated the orbit using \texttt{gala} and
the \texttt{MWPotential2014} potential \citep{bovy_galpy_2015,gala}.

The general shape of the cluster itself seems to include a central overdensity
and two tails (the halo).  One tail is leading the cluster's orbit,
and is angled toward the center of the galaxy when viewed top-down.
The other tail is trailing the cluster's orbit, and is pointed away
from the center of the galaxy.  The elongation of the cluster in both
the $(X,Y)$ and $(Z,Y)$ planes is correlated with the direction of the
LSR-corrected median cluster velocity.   Possible explanations for
this overall morphology are discussed in Section~\ref{subsec:origin}.

@@

The first check on whether the candidate cluster members
share the same age is to analyze the Gaia Hertzsprung-Russell (HR)
diagrams.  Comparable analyses have previously been performed by
\citetalias{cantatgaudin_gaia_2018},
\citetalias{kounkel_untangling_2019}, and \citetalias{meingast_2021}.

Figure~\ref{fig:hr} shows the HR diagram in the space of absolute Gaia
$G$ magnitude as a function of observed $Bp-Rp$ color.  Stars in
the core appear consistent with
having a fixed age and metallicity, and varying mass.  The halo stars
show a similar sequence, though with greater scatter.  One possible
explanation for the additional scatter could be that the halo is more
contaminated by field stars.  For instance, $\approx5$ red
giants in the halo
must be field interlopers, because their isochronally implied
ages would be inconsistent with that of the cluster.

@@

To determine the reddening correction across \cn, we
queried the 2MASS
DUST
service\footnote{
	\url{http://irsa.ipac.caltech.edu/applications/DUST};
	query performed using the \texttt{astrobase.services.dust} module
	\citep{bhatti_astrobase_2018}.
}
and retrieved the extinction parameters at the positions of each
NGC\,2516 member.  The mean and standard
deviation of the $E(B-V)$ values for the \citet{schlegel_maps_1998}
and \citet{schlafly_measuring_2011} maps were consistent, so we
adopted the average from \citet{schlegel_maps_1998}: $E(B-V)_{\rm
S}=0.206\pm0.039$.  \citet{bonifacio_search_2000} found however that
the \citet{schlegel_maps_1998} maps overestimate the reddening values
when the color excess exceeds about 0.10\,mag. Applying the correction
proposed by \citet{bonifacio_search_2000} leads to our adopted
$E(B-V)=0.102\pm0.019$.  Finally, to convert to Gaia colors we used
the calibration of \citet{stassun_TIC8_2019}, namely $E(Bp-Rp)=1.31
E(B-V)$ and $A_G=2.72 E(B-V)$.  This yielded $E(Bp-Rp)=0.134\pm0.025$,
which is used in the plots.  To redden the isochrones, we assumed
$R_V=3.1$, and applied the \citet{odonnell_1994} SED-dependent
extinction law star-by-star through the PARSEC interface.

@@

Overall, the data and models agree for masses above
$\approx$0.45\,$M_\odot$.  Below this mass, the data and models diverge at
colors redder than $Bp-Rp\approx2.2$, in the sense that the model
isochrones have lower luminosities and bluer colors than the
observations.\footnote{The 100 Myr isochrone does overlap well with the
lower main sequence, but it fails to fit the upper main sequence and
red giants.}  The MIST isochrones showed a comparable disagreement
\citep{choi_mesa_2016}.  Proposed explanations for the discrepancy
between the models and observations have include starspots and
incomplete molecular line lists\citep[{\it
e.g.},][]{stauffer_why_2003,feiden_magnetic_2013,rajpurohit_effective_2013,mann_spectrothermometry_2013,choi_mesa_2016}.
Ultimately, we adopted the PARSEC isochrones because
\citet{chen_improving_2014} implemented an empirical
temperature-opacity calibration, which leads the PARSEC isochrones to
diverge at slightly lower mass than the MIST models.  Regardless, for
purposes of age-dating the cluster we focus on the main-sequence
turn-off (MSTO), since this is where the models have maximal
predictive power.

\citet{cummings_2018} presented techniques for mitigating some of the
complexities of MSTO age-dating ({\it e.g.}, sparse turnoffs, stellar
rotation, high binarity fractions, and the presence of blue
stragglers).  Combining a $UBV$ color-color analysis with Gaia DR2
cluster memberships, they found MSTO ages for NGC\,2516 ranging from
165 to 195\,Myr, depending on the choice of model isochrone (Y$^2$,
PARSEC, MIST, or SYCLIST).  %Their statistical uncertainties ($\approx
%25$\,Myr) were comparable to the systematic uncertainties.

Our goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
isochronally, they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim
can be made.  The MSTO stars are consistent with the 178 and 316 Myr
models, but are bluer than the 100 Myr model.  The red-giant-branch (RGB) stars are
consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
model.  Based on the assumption that the width of the MSTO can be
attributed to binary stars, we are most interested in the blue edge
(blue stragglers are, however, a concern).  The blue edge appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for the more precise model-dependent
comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
adopt a PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
our quoted uncertainty is a quadrature sum of the statistical
and systematic components from \cite{cummings_2018}. 
%The relative age of ``slightly older than
%the Pleiades'' is likely the more robust statement, consistent with
%reports since \citet{mermilliod_comparative_1981}.

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo successive stages of automated cleaning,
    (see Section~\ref{subsubsec:cluster}).  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    From $1.2<(Bp-Rp)_0<1.7$, 
    some NGC\,2516 members have longer rotation periods than Pleaides members.  
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

We began our rotation analysis by considering all \nkinematic\
candidate members of \cn.  For each source, we retrieved all available
CDIPS light curves, on a per-sector basis.  This yielded \ncdips\
stars with at least one sector from a CDIPS light curve.  Each of
these stars is brighter than the CDIPS magnitude limit of
$Rp=16$\,mag.  \nkinematicrpltsixteen\ of the stars from the \cn\
source list had $Rp<16$.  The difference (65 stars)
is due to 35 stars from 
\citet{meingast_2021} which were not available at the time of the
CDIPS reductions, and 30 stars falling on the TESS chip gaps.  The
$Rp=16$\,mag cutoff imposed during the CDIPS processing corresponds
roughly to $(Bp-Rp)_0$ of 2.2, or a spectral type of $\approx$M2V, at
the distance of \cn.

We removed systematic trends shared between all light curves on each
CCD in each sector individually, and stitched together the resulting
light curves before searching for rotation-induced periodicity.
Details regarding our detrending approach are presented in
Appendix~\ref{app:detrending}.  After applying the initial
ensemble decorrelation to remove systematic trends, we proceeded with a few
cleaning steps: we masked 0.7 days at the beginning and end of each
spacecraft orbit, and ran a sliding standard-deviation rejection
window over the light curve, which removed any outlying points within
$\pm3\times$MAD of the median in each window.

We then measured the rotation period from the resulting light curve
using the periodogram implementations in \texttt{astrobase}, namely
the traditional Lomb-Scargle, and
the Stellingwerf phase-dispersion minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light
from the target star and background-light \citep{Sullivan_2015}.  This
typically resulted in an aperture radius of either 1 or 1.5 pixels.
We recorded the top five periodogram peaks from each method, and their
corresponding powers.  Finally, as a check on crowding, we recorded
the number of stars within the aperture with equal or greater
brightness than the target star, and with brightness within 1.25 and
2.5 TESS magnitudes of the target star.
